### 2022-02-08 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd
Update chan_cfrd_metabolomics.Rmd
PLSDA permutation tests

```{r setup, include=FALSE}
library(Hmisc)
library(mixOmics)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
plsda_res = splsda(X, Y_all)
plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16,
          title = "PCA Components 1 - 2")
```

>>>>

```{r setup, include=FALSE}
library(Hmisc)
library(mixOmics)
library(combinat)
library(tidyverse)
library(parallel)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
plsda_res = splsda(X, Y_all)
plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16)
```

```{r}
plsda_res = splsda(X, Y_two)
plotIndiv(plsda_res,group = df$combined,legend = TRUE,pch = 16)
```

### Permutation testing

```{r cache=TRUE}
real_auc_1 = auroc(plsda_res,print = F,plot = F)$Comp1[1]
real_auc_2 = auroc(plsda_res,print = F,plot = F)$Comp2[1]
perms = 1000
cl = makeCluster(8,type = "FORK")
aucs = parLapply(cl,1:perms, function(s){
  set.seed(s)
  Y_perm = sample(df$combined,length(df$combined))
  plsda_res = splsda(X, Y_perm)
  aucs = auroc(plsda_res,print = F,plot = F)
  comp1_auc = aucs$Comp1[1]
  comp2_auc = aucs$Comp2[1]
  return(c(comp1 = comp1_auc,comp2 = comp2_auc))
})
stopCluster(cl)
aucs = as.data.frame(do.call(rbind,aucs))
```

Two group outcome was "shuffled" `r perms` times and sPLSDA re-fit on the permuted outcome.

#### Component 1

```{r}
ggplot(aucs,aes(x = comp1)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
  geom_vline(xintercept = real_auc_1,color = "red") +
  xlab("AUC") + ylab("Count")
```

Permutation p value for component 1 is `r round((1 + sum(aucs$comp1 > real_auc_1))/perms,3)`

#### Component 2

```{r}
ggplot(aucs,aes(x = comp2)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
  geom_vline(xintercept = real_auc_2,color = "red") +
  xlab("AUC") + ylab("Count")
```

Permutation p value for component 1 is `r round((1 + sum(aucs$comp2 > real_auc_2))/perms,3)`
